Eagle: Making multiple-locus association mapping on a genome-wide scale routine

### 2018-08-01 paper.tex
1August refining v3

We purposely chose  methods that differed from one another. BigRR, LMM-Lasso, and glmnet are all regression-based regularisation 
methods. However, BiggRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
They allow the strength of association between all the SNPs and trait to be measured within a single model, simultaneously. 
A limitation of these methods though is that the statistical significance of the SNP effects cannot be easily determined. 
 Due to the adaptive nature of the estimation procedures, to do this 
analytically is challenging and is an area of active research \ref{lockhart2014significance}.

@@

R2VIM is different to the other four methods in that it is a non-parametric model-free approach. It implements  
random forests and a new way of measuring the importance of a SNP.  
Here, multiple parallel runs are performed. Within a run, a relative importance score is calculated for each SNP and 
calculated by dividing the importance score for a SNP by a run's minimum importance score. 
Only those SNPs with relative importance scores above a certain threshold across all the runs are deemed to be significant. 
The relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
but the computation cost is high.

 The greater the score is than one, 
the stronger the evidence is that the SNP is in association with the trait.  Only those SNPs with scores above a certain threshold are 
deemed as being significant. The difficulty is that the relationship between this threshold and false positive rate is unknown. The threshold could be 
found empirically, via permutation but this is computationally demanding.

In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. 
It is from these importance scores that SNPs can be ordered in terms of their strength of association with a trait. 
The difficulty is in knowing what proportion of the 
highest ordered SNPs are in true association. R2VIM addresses this by performing multiple parallel runs. Each run uses a different 
random seed, giving slightly different results for the random forest analyses. Within a run, a relative importance score ($f$) is calculated for 
each SNP. The greater $f$ is than one, the more evidence there is that the SNP is in association with the trait.  A SNP-trait association is 
declared if $f$ for the SNP is above some threshold value across all parallel runs. The difficulty is that the relationship between 
this threshold value and the chance of finding false positives is unknown.  The threshold value could be found via permutation but 
this would be computationally challenging for all but the smallest of association studies.

@@

GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast} are of  the same computational complexity \cite{zhou2012genome}, produce exact instead of approximate results, and are highly efficient computationally.  Both perform a single spectral decomposition of the relationship matrix $K$, use  eigenvector matrix to rotate the data, 
and reformulate the (reml) log likelihood for easier computation.  They differ though in their estimation procedure. 
FaST-LMM implement's the Brent's algorithm. GEMMA instead implement's the Newton-Raphson algorithm. The  Newton-Raphson is more complicated  but has superior 
convergence properties.

===>

log likelihood into a sum of $n$ terms that are easier to compute. The difference lies in their estimation procedure. FaST-LMM implement's the Brent's algorithm to optimise $\delta$. GEMMA instead implement's the Newton-Raphson algorithm. Newton-Raphson is more complicated in that it also requires the first and second derivatives of a function to be calculated. However, it is superior in terms of its convergence properties to the Brent algorithm.  Both applications are stand-alone computer programs, popular, and in common use.

%%\fi

>>>>

We compared the computational and statistical performance of Eagle against five multiple-locus methods. 
BigRR, LMM-Lasso, and glmnet are all regression-based regularisation 
methods. BiggRR is based on generalised ridge regression, LMM-Lasso is based on lasso, and glmnet is based on elastic net. 
Regularisation methods make parameter estimation possible in models  where the number of predictors is far greater than the number of samples. 
They allow the strength of association between all the SNPs and trait to be measured within a single model, simultaneously. 
A limitation of these methods though is that the statistical significance of the SNP effects cannot be easily determined. 
 Due to the adaptive nature of the estimation procedures, to do this 
analytically is challenging and is an area of active research \ref{lockhart2014significance}. Instead, we calculate significance empirically via 
stability selection (see below).

@@

R2VIM differs to the other four methods in that it is a non-parametric model-free approach. It implements  
random forests but where multiple parallel runs are performed. Each run begins with a different random number seed, leading to 
different random forests being formed.  A relative importance score is calculated, within a run,  for each SNP. This is done by dividing a SNP's 
importance score by the minimum importance score observed across all the SNPs within a run. 
Only those SNPs with relative importance scores above a certain threshold across all the runs are deemed to be significant. 
Unfortunately, the relationship between threshold value and false positive rate is unknown. The threshold could be found empirically via permutation 
but the computational cost is high, restricting the size of data that can be analysed.

@@

GEMMA \cite{zhou2012genome} and FaST-LMM \cite{lippert2011fast} are of  the same computational complexity \cite{zhou2012genome}, produce exact instead of approximate results, and are highly efficient computationally.  Both perform a single spectral decomposition of the relationship matrix $K$, use  eigenvector matrix to rotate the data, 
and reformulate the (reml) log likelihood for easier computation.  They differ in their estimation procedure. 
FaST-LMM implement's the Brent's algorithm. GEMMA instead implement's the Newton-Raphson algorithm. The  Newton-Raphson is more complicated  but has superior convergence properties to Brent's algorithm. 
Both applications are stand-alone computer programs and in common use.
